---
title: "In Class Lab"
author: "Emanuel Padilla"
date: "DATA 2401"
output: html_document
---
**Note: Your exam will be in a similar format to this (RMarkdown File with Questions and Code chunks for your answers), so make sure you have the process for knitting to html working.** 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In-Class Exercises

### Set 1 

1. Use the colon operator : to create a vector `n` of numbers from 10:49

```{r}
n <- 10:49

```

2. Use the `length()` function to get the number of elements in `n`
```{r}
length(n)

```

3. Add 1 to each element in `n` and print the result
```{r}
n_result <- paste(n, 1, "")

n_plus_1 <- n + 1
```
4. Create a vector `m` that contains the numbers 10 to 1 (in that order), in two different ways, neither of which is just typing "10, 9 , 8 ...".  Hint: use one of the helper functions we learned.
```{r}
m <- 10:1

# m <- seq(10, 1)
```
5. Subtract `m` FROM `n`. Note the recycling!
```{r}
vec <- n - m
vec


```
### Set 2 

1. Use the `seq()` function to produce a range of numbers from -5 to 10 in `0.1` increments. Store it in a variable `x_range`
```{r}
x_range <- seq(-5, 10, by = 0.1)
```
2. Create a vector `sin_wave` by calling the `sin()` function on each element in `x_range`.
```{r}
sin_wave <- sin(x_range)
```
3. Use the `plot()` function to plot your "wave"!
     + the x vector goes in the first argument, the y vector goes in the second argument
     + You'll need to add the argument `type = "l"` for a line. NOTE: That is a lower case "L", not a 1. 
```{r}
plot(x_range, sin_wave, type = "l")
```


### Set 3 

1. Use `:` to make a vector of the integers from 40 to 50 called `first_nums`
```{r}
first_nums <- 40:50

```
2. Use `seq()` to make a vector of the integers from 51 to 60 called `next_nums`
```{r}
next_nums <- seq(51, 60)
```
3. Combine them using `c()`
```{r}
comb_vec <- c(first_nums, next_nums)
```
4. Use indexing to access the 7th number in your combined vector
```{r}
comb_vec[7] 
```
5. Use indexing to access the 5th to the 8th numbers in your combined vector
```{r}
comb_vec[5:7]
```
6. Use indexing to create a new vector that contains the first 3 elements and last 3 elements of your combined vector. 
```{r}
first_three<- comb_vec[1:3]

length <- length(comb_vec)

last_three <- comb_vec[(length - 2):length]


first_n_last <-c(first_three, last_three)
first_n_last
```
7. Bonus/extra: do the above in a way that would work no matter how long the vector is (as long as it's longer than 2 elements)
```{r}
n <- length(comb_vec)
first_n_last <- c(comb_vec[1:3], comb_vec[(n-2):n])
first_n_last
```

## In-Class Activities

1. Missingness 

* Use is.na(x) to find which entries are missing in the vector below.
* Create x_no_na that removes the missing values.
* Replace the missing values with the mean of the non-missing values (compute the mean yourself from x_no_na).
     * Hint: x[is.na(x)] <- ...

```{r}
x <- c(3.2, NA, 4.1, 5.0, NA, 2.8)


#removes the missing values
x_no_na <- x[!is.na(x)]

#gets the mean from vectors without the na
mean_x <- mean(x_no_na)

#replaces the na with the mean
x[is.na(x)] <- mean_x
x


```

1. Logical indexing

* Using the `grades` vector from class:
* Create a logical vector is_a <- grades >= 90
* Print the values that are A’s (you already do this)
* Now print the indices/positions of the A’s using which(is_a)
* How many A’s are there? Do it two ways:
     * length(grades[is_a])
     * sum(is_a) (since TRUE counts as 1)

```{r}
grades <- c(99,  77, 100,  91,  87,  72,  82,  94,  81)

is_a <- grades >= 90

#prints the values that are A's
grades[is_a]

#prints indices of A's
which(is_a)

#Both gets us how many A's
length(grades[is_a])
sum(is_a)
```

1.  Emperical probabilities

* Set a seed: `set.seed(2401)`
* Simulate 200 die rolls (fair die) in a vector rolls (use sample(1:6, ...))
* Use table(rolls) to count how many of each face occurred.
* Convert counts to proportions (estimated probabilities) by dividing by length(rolls).
* Which face was most common in your simulation?


```{r}
set.seed(2401)

rolls <-sample(1:6, 200, replace = T)

#counts how many of each face occurred
table(rolls)

#counts to estimated probabilities
est_prob <- table(rolls) / length(rolls)
est_prob

# The most common face in my simulation was 4.

```


1. Write a *function* to simulate rolling a 6-sided die and returns the result.

```{r}
dice <- function(rolls){
  rolling <- sample(1:6, 1, replace = T)
  rolling
}

dice(1)
```





1. Write a function to simulate rolling 2 6-sided dice that returns the sum of the rolls. 

```{r}
total_rolls_with_sum <- function(rolls_){
  
rolling_ <- sample(1:6, rolls_, replace = T)

sum_rolling <- sum(rolling_)

print(rolling_); print(sum_rolling)
}

total_rolls_with_sum(2)

    
  
 

```


1. Write a function that takes two people name's as strings, simulates rolling a 6-sided die for each of the names, then returns the name of the person with the higher roll. The function should reroll once on a tie, but if there is a second tie, it should return "draw". 

```{r}

two_names_two_rolls <- function(name_1, name_2){
  
#CREATE A VECTOR FOR THE NAMES
names <- c(name_1, name_2)

#ROLLS THE 6SIDED DICE 2 TIMES
rolling <- sample(1:6, 2, replace = T)

#highest roll
max <- which.max(rolling)

#RETURNS the NAME WITH THE highest ROLL, RESPECTIVE TO THE HIGHEST ROLL INDECE
name_max <- paste(names[max], rolling[max])

#RETURNS BOTH NAMES WITH THE ROLL RESPECTIVELY
sum_ <- paste(names, rolling)

#IF THE FIRST ROLL DOES NOT HAVE A TIE, DONE
  if(rolling[1] != rolling[2]){
     print(name_max);print(sum_);
 
#ELSE THE FIRST ROLL DOES HAVE A TIE, REDO THE ROLL
  }else{    

    rolling2 <- sample(1:6, 2, replace = T)
    max2 <- which.max(rolling2)
    name_max2 <- paste(names[max2], rolling2[max2])

    #REDO: IS THERES NO TIE NOW, DONE
    if(rolling2[1] != rolling2[2]){
      sum_2 <- paste(names, rolling2)
      print(name_max2); print(sum_2)
      
    }else{    #REDO: THERES A TIE AGAIN, DRAW
      print("DRAW")
    }
  }
}

two_names_two_rolls("Emanuel","Abel")

```


1. Write a second function (you may start by modifying the first one) that takes two names as strings, simulates both of them rolling die 15 times, and returns the name of the person who "won" more rolls. Don't reroll ties this time, but don't count them as "wins". 

Process: 

* Simulate 15 rolls each for two players, p1 and p2 (vectors).
* Create a logical vector p1_wins <- p1 > p2
* Create a logical vector p2_wins <- p2 > p1
* Create a logical vector ties <- p1 == p2
* Report:
     * number of p1 wins (sum(p1_wins))
     * number of p2 wins (sum(p2_wins))
     * number of ties (sum(ties))

```{r}
roll_15_ <- function(name_1, name_2){
  

#ROLLS THE 6SIDED DICE 15 TIMES
p1_wins <- sample(1:6, 15, replace = T)
p2_wins <- sample(1:6, 15, replace = T)

total_p1 <- sum(p1_wins)
total_p2 <- sum(p2_wins)

if(total_p1 > total_p2){
  print(name_1)
} else if(total_p2 > total_p1){
  print(name_2)
}else{
  print("TIE! NO WINNER!")
}
}

roll_15_("Steve","Miguel")

```


3. Modify your function from (2) so that the person in the first argument has a die that is weighted to get higher numbers more often (you may decide the actual percentage change). Run the function ten or so times. Does the second person ever win?

In my case, player 2 did not win the ten times. I mega cheated with player ones dice be weighted to roll 5s and 6s. 

```{r}
cheating_roll <- function(name_1, name_2){
  

#ROLLS THE 6SIDED DICE 15 TIMES, p1 IS WEIGHTED
p1_wins <- sample(5:6, 15, replace = T)
p2_wins <- sample(1:6, 15, replace = T)

total_p1 <- sum(p1_wins)
total_p2 <- sum(p2_wins)

if(total_p1 > total_p2){
  print(name_1)
} else if(total_p2 > total_p1){
  print(name_2)
}else{
  print("TIE! NO WINNER!")
}
}

cheating_roll("Sofia","Margaret")

```

